---
title: "lab01-josh-reinheimer"
author: "josh reinheimer"
date: "15 Februar 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting started
The purpose of this notebook will be threefold: First, I will be trying to gain more experience in writing R markup documents. Second, I will be familiarizing myself with the R programming language and RStudio. Thirdly, I will be working through the lab and completing all small assignments that are given.


### Your Turn # 1
```{r first constant}
cellphone <- 80
transportation <-  20
groceries <- 527
gym <- 10
rent <- 1500
other <- 83
total <- cellphone + transportation + groceries + gym + rent + other
total
semester = total * 5
year = semester * 2
semester
total
```

### Your turn #2
```r{second}
expenses = c(cellphone,transportation,groceries, gym, rent, other)
sorted_expenses <- sort(expenses, decreasing = TRUE)
xx <- barplot(sorted_expenses,names.arg = c("rent", "groceries", "other", "cell", "trans.", "gym"),cex.names=.5)
text(x = xx, y = sorted_expenses, label = sorted_expenses, cex = 0.8, col = "red")
```
### Your turn #3
```r{third}
hypotenuse =  sqrt(3^2 + 4^2)
print(hypotenuse)
```
### Your turn # 4
```r{fourth}
#considering a fair coin
n <- 5
k <- 2
p <- .5
nchoosek <-  factorial(n)/(factorial(k) * factorial(n-k))
print(nchoosek)
binomial_factorial <-  nchoosek*(p^n)
print(binomial_factorial)
binomial_choose <- choose(n,k)*(p^n)
print(binomial_choose)

#considering a fair coin
n_d <-10
k_d <- 3
p_d <- .167

binomial_die = choose(n_d,k_d)*(p_d^k_d)*((1-p_d)^(n_d-k_d))
print(binomial_die)
# 2 heads in 5 fair coin tosses
print(dbinom(2,5,.5))
# 3 6s in 3 fair die tosses
print(dbinom(3,3,.167))
# more than 3 heads in 5 tosses of a biased coin (35 % chance of heads)
print(sum(dbinom(4:5,5,.35)))
```
### Your turn #5
```r{fifth}
install.packages(c("stringr", "RColorBrewer", "XML"))


x <-  c(-3:3)
print(3*x^2 + 4*x + 8)

```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
