---
title: 'lab-11: Computing Probabilities'
author: "Josh Reinheimer"
date: "6 März 2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Random number generators are used in a wide range of scientific and mathemaical applications. Anytime there is uncertainty present in a problem, it can be modeled via random numbers
## Computing Probabilities
if we were to have two bags containing a number of different balls, we can use R to simulate the experiment of picking a bag and drawing a ball. To accomplish this, we would first create two vectors, which would simulate the bags.

```{r}
bag1 <- c('white', 'white', 'red')
bag2 <- c(rep('white', 3), 'red')
```
The experiment of drawing will then be repeated a number of times. The drawing will take place with replacement. 

```{r}
bags <- c('bag1', 'bag2')
repetitions <- 1000
drawn_balls <- character(repetitions)

set.seed(345)
for (i in 1:repetitions) {
  # select one bag
  chosen_bag <- sample(bags, 1)
  
  # draw a ball from chosen bag
  if (chosen_bag == 'bag1') {
    drawn_balls[i] <- sample(bag1, 1)
  } else {
    drawn_balls[i] <- sample(bag2, 1)
  }
}

table(drawn_balls) / repetitions
```
This problem can be made more complicated by considered two different boxes filled with balls of different colors. The first contains two blue balls and one red ball. The second two blue, three red, and one white

```{r}
box1 <- c(rep('blue',2), 'red')
box2 <- c(rep('blue',2), rep('red', 3), 'white')
```

The drawing rules are then established. If a randomly generated number is greater than .5, then a sample with replacement of size 4 is drawn from box 1. If the number is equal to or smaller than .5, then a sample without replacement of size 4 is drawn from box 2. 
```{r}
set.seed(2345)
n <- 1000
drawn_balls <- matrix(0,nrow = n,ncol = 4)

for(i in 1:n){
  if(runif(1) > 0.5){
    drawn_balls[i,] <- sample(box1, size = 4, replace = TRUE)
  } else{
    drawn_balls[i,] <- sample(box2, size = 4, replace = FALSE)
  }
}
```
Once this is repeated 1000 times, we are interested in the frequencies of the different number of blue balls, namely 0, 1, 2, 3, and 4. The plots below show the relative frequencies of the number of blue balls drawn
```{r}
occurences <- rep(0,n)
for(i in 1:n){
occurences[i] <- sum((drawn_balls[i,]=="blue")==TRUE)
}
freq_0 <- cumsum(occurences == 0) / 1:n
freq_1 <- cumsum(occurences == 1) / 1:n
freq_2 <- cumsum(occurences == 2) / 1:n
freq_3 <- cumsum(occurences == 3) / 1:n
freq_4 <- cumsum(occurences == 4) / 1:n
n_series = seq(1:n)
frequencies <- data.frame(n_series,"0" = freq_0,"1" = freq_1, "2" = freq_2,"3" = freq_3,"4" = freq_4)
```
Plotting the relative frequencies of the different numbers of blue balls, we find
```{r, echo = FALSE}
library(ggplot2)
ggplot(frequencies, aes(n_series, y = value, color = number)) +
  geom_line(aes(y = freq_0, col = "0")) +
  geom_line(aes(y = freq_1, col = "1")) +
  geom_line(aes(y = freq_2, col = "2")) +
  geom_line(aes(y = freq_3, col = "3")) +
  geom_line(aes(y = freq_4, col = "4")) +
  ggtitle("Relative frequencies of the number of blue balls") +
  xlab("reps") + ylab("freq.") 
```