Much of what we do in programming requires performing some task over and over again. This particular behavior is called a loop. 

In the R language, the majority of functions are vectorized, which means that will apply to every element in a vector. In this lab, we will act as if R does not have vectorized operations and will instead build loops to perform the desired operations

## For loop
The basic basic type of loop is called a for loop. In R, a for loop is defined as
```{r}
#for(i in start:end){
  #do something
#}
```
i is called the counter of the loop. The counter is initialized to the value that is given in start and after each execution of the loop is incremented by a value. Unless otherwise specified, this value is 1. The process continues until the counter equals the value of end. 

Included below are two small for loops that demonstrate how they work

```{r}
#this loop multiples each value of the vector vec by three
vec <- c(3,1,4)
for(j in 1:length(vec)){
  vec[j] <- vec[j]*3
}

# this loop does the same thing, but saves the results to a different vector
vec2 <- rep(0,length(vec))
for(i in 1:length(vec)){
  vec2[j] <- vec[j]*3
}
```
## Summation series
One of the most important applications of for loops is for the mathematical process of summation. In the following code snippets, we have coded different kinds of summation series 
### arithmetic series
```{r}
total = 0
n = 100
a1 = 3
d = 3
series = rep(0,n)
for(i in 1:n){
  total = a1 + (i-1)*d
  series[i] = total
}
print(series)

```
This series does not converge.
### geometric series
```{r}
n = 20
geom = rep(0,n)
a1 = 3
r = 2
for( i in 1:n){
  geom[i] <- a1*r^(i-1)
}
sum(geom)
```
### sine approximation
```{r}
point = 10
total = 0
n = 100
for( i in 1:n){
  total = total + (point^((2*i)-1)/factorial((2*i)-1))*((-1)^(i-1))
  }
print(total)
```
## For loop with a matrix
For computational problems in applied mathematics, the ability to manipulate matrices is of upmost importance. In fact, any in depth problem will almost certainly involve matrices at some point. Using loops, one can perform many different matrix operations

In the following code, we start with a matrix of randomly select values. The negative values in the matrix will be transformed into squared values, whil the positive values are transformed into square root values.
```{r}
X <- matrix(rnorm(12), nrow = 4, ncol = 3)
print(X)
for(i in 1:nrow(X)){
  for(j in 1:ncol(X)){
    if(X[i,j] >= 0){
      X[i,j] <-  sqrt(X[i,j])
    }
    else{
      X[i,j]  <-  (X[i,j])^2
    }
  }
}
print(X)
```
Sometimes one wishes to continue a loop until a particular condition is met. In this case, a while loop is used. A while loop has the following form
```{r}
#while(statement == TRUE){
  # do something
#}
```
As long as our statemnt in question evaluates as TRUE, the loop will continue running. As soon as it evaluates FALSE, then it will stop.
Using a while loop, we will create a function that will continuously divide a number by two until it becomes odd
```{r}
#' @title Reduction function
#' @description The scalar value will be continuously
#' reduced until the value becomes odd. This odd value 
#' will then be returned
#' @param x. Even numerical value (numeric)
#' @return y. Odd numerical value (numeric)

reduce <- function(x){
  while(x%%2 == 0){
    x <-  x/2
  }
  return(x)
}
value_while <- 898128000
print(reduce(value_while))
```
Using for loops, while loops, and repeats we will create the average, the standard deviation(SD), and geometric mean of a vector of values from 1 to 50
```{r}
##### average #####
vector = seq(1,100,1)
mean_for = 0
mean_while = 0
mean_repeat = 0
for(i in 1:length(vector)){
  mean_for = mean_for + vector[i]
}
for_mean = mean_for/length(vector)
cat("for loop mean: ", for_mean)
temp = 0
for(i in 1:length(vector)){
  temp = temp + (vector[i] - for_mean)^2
}
cat("for loop SD ", sqrt((1/(length(vector)-1))*temp))

i = 1
while(i <= length(vector)){
  mean_while <- mean_while + vector[i]
  i <- i + 1
}
while_mean = mean_while/length(vector)
cat("while loop: ", while_mean)

temp = 0
i = 1
while(i <= length(vector)){
  temp <- temp + (vector[i] - while_mean)^2
  i <- i + 1
  }
cat("while loop SD ", sqrt((1/(length(vector)-1))*temp))



j = 1
repeat{
  if(j < 101){
    mean_repeat <- mean_repeat + vector[j]
    j <- j + 1
  }
  else{
    break
  }
}
repeat_mean <- mean_repeat/length(vector)
cat("repeat loop mean: ", repeat_mean)

j = 1
temp = 0
repeat{
  if(j < 101){
    temp <- temp + (vector[j] - repeat_mean)^2
    j <- j + 1
  }
  else{
    break
  }
}
cat("repeat loop SD: ",sqrt((1/(length(vector)-1))*temp))

##### Geometric Mean #####
for_geo_mean = 1
vector = seq(1,50,1)
for(i in 1:length(vector)){
  for_geo_mean <-  for_geo_mean*vector[i]
}
for_geo_mean <- for_geo_mean^(1/length(vector))

while_geo_mean <-  1
i = 1
while(i <= length(vector)){
  while_geo_mean <- while_geo_mean * vector[i]
  i <- i + 1
}
while_geo_mean <- while_geo_mean^(1/length(vector))

repeat_geo_mean <- 1
j <- 1
repeat{if(j <= length(vector)){
  repeat_geo_mean <- repeat_geo_mean*vector[j]
  j <- j + 1
}
  else{
    break
  }
}
repeat_geo_mean <- (repeat_geo_mean)^(1/length(vector))

```

### Distance matrix of letters
Combining loops and matrices, we will look at a problem of a matrix consisting of arbitrary distance values amount letters in the English language
```{r}
num_letters <- length(LETTERS)
set.seed(123)
values <- sample.int(num_letters) 
distances <- values %*% t(values)
diag(distances) <- 0
dimnames(distances) <- list(LETTERS, LETTERS)
print(distances[1:5,1:5])
```
From this distance matrix, one can calculate the distance total of a word. Using the vvalues above, the vector ('E', 'D', 'A') has a total score of 651. The distance from 'E' to 'D' is 483, while the distance from 'D' to 'A' is 168. 
With this in mind, we will calculate the scores for a number of different vectors stored in a list entitled 'string'.

```{r}
get_distance <- function(distance, letter_vec){
  for(i in 1:length(letter_vec)){
    if(letter_vec[i] %in% LETTERS == TRUE){
      
    }else{
      stop("Element is not a letter")
    }
  }
  sum = 0
  for(i in 1:(length(letter_vec)-1)){
    sum <- sum + distances[letter_vec[i],letter_vec[i+1]]
  }
  return(sum) 
}
cal <- c('C', 'A', 'L')
stats <- c('S', 'T', 'A', 'T', 'S')
oski <- c('O', 'S', 'K', 'I')
zzz <- rep('Z',3)
lets <- LETTERS
first <- c('J', 'O', 'S', 'H')
last <- c('R', 'E', 'I', 'N')
strings <- list(
  cal = cal,
  stats = stats,
  oski = oski,
  zzz = zzz,
  lets = lets, 
  first = first,
  last = last
)
strings_dist <- rep(0,length(strings))
counter = 1
for(i in strings){
  strings_dist[counter] <- get_distance(distances,i)
  counter <- counter + 1
  }
print(strings_dist)
```
