---
title: "warmup03-josh-reinheimer"
author: "Josh Reinheimer"
date: "20 Februar 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Create the dataset. The condition header = TRUE transforms the first column into the header names
github <- 'https://github.com/ucb-stat133/stat133-fall-2018/'
repo <- 'raw/master/data/saratoga.txt'
house <- read.table(paste0(github, repo), sep = "\t", header = TRUE)
libaray(plyr)


# Summary of column Price
summary(house$Price)

# Histogram of the housing prices
hist(house$Price)

# Log transformation of prices and ccreation of a new column
# for these transformed values
house$log_Price <- log(house$Price)
summary(house$log_Price)
hist(house$log_Price)

# Determining the number of houses that are "cheap"
cheap <- house$log_Price< 10.5
length(cheap)
# Drop the rows(houses) that appear in the array "cheap"
house <- house[!cheap,,drop = TRUE]

# Explore the Living Area
summary(house$Living.Area)
hist(house$Living.Area)

# Transform the Living Area values
house$log_Living.Area <- log(house$Living.Area)
# The value of approximately 6.8 in log_Living.Area corresponds to approximately 800-1000 sq. feet
# in the column Living.Area

# Here we will transfrom the column Bedrooms
# into a categorical data, containing the values
# <2, 3, 4, and >5
BR <- house$Bedrooms
# Replacing values <2 or > 5 with the corresponding strings
BR[BR > 5] = 5
BR[BR < 2] = 2
factor_BR <- factor(BR, level = c(2, 3, 4, 5), labels = c("<2", "3", "4", ">5"), ordered = TRUE)
summary(factor_BR)

# Plot of the price against living area
my_colors <- c('aquamarine3', 'darkgoldenrod2', 'coral2', 'mediumorchid3')
br_colors <- mapvalues(factor_BR, c("<2", "3", "4", ">5"), my_colors) 

plot( Price ~ Living.Area, data = house, log = "xy", main = "",
      xlab = "Living Area (log sq ft)", ylab = "Price (log $)",
      col = br_colors, pch = 19, cex = 0.4)
legend("bottomright", fill = my_colors, legend = levels(factor_BR),
       title = "# Bedrooms", cex = 0.75)
# Number of fireplaces
# First we transform the number of fireplaces 
# into categorical data. the value 0 is replaced with None
# and any value greater than 0 with At least 1
FP <- house$Fireplaces
FP[FP > 0] = 'At least 1'
FP[FP == 0] = 'None'
factor_FP <- factor(FP)
summary(factor_FP)
house$factor_FP <- factor_FP
house$PPSF <- house$Price/house$Living.Area

my_colors1 <- c('aquamarine3', 'darkgoldenrod2')
fp_colors <- mapvalues(factor_FP, c('None', 'At least 1'), my_colors1) 


plot(house$Price, house$PPSF,log = "xy", main = "",
     xlab = "Price ($)", ylab = "Price per square foot ($)",
     col = fp_colors, pch = 19, cex = 0.4)
legend("bottomright", fill = my_colors1, legend = levels(factor_FP),
       title = "# of fireplaces", cex = 0.75)






```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
