---
title: "warmup-6-josh-reinheimer"
author: "Josh Reinheimer"
date: "11 März 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```
# Warm up 6: Control Flow Structures

## Introduction
Using loop structures, in combination with functions, allows one to solve a number of different problems and generalize the techniques developed to a greater number of problems.

One of the most well studied probability distributions is the binomial distribution. This is a probability model that allows one to calculate the number of successes in a fixed number of trials. If one is tossing a coin, questions arise about the likelihood of certain sequences. What is, for example, the likelihood of getting three heads in four tosses of a fair coin, or P(X 0 3 heads in four tosses). This can be calculated by the following formula:
$$Pr(X = k) = \binom{n}{k}p^{k}(1-p)^{n-k} $$
where:
  * n is the number of (fixed) trials
  * p is the probability of success on each trial
  * 1 ??? p is the probability of failure on each trial
  * k is a variable that represents the number of successes out of n trials
  * the first term in parenthesis is not a fraction, it is the number of combinations in which k success can occur in n trials
  
So for 3 heads, we can calculate
$$Pr(X = 3) = \binom{4}{3}0.5^{3}(1-0.5)^{4-3} = .25 $$

In this warm up, functions and loop structures will be used to build the binomial distribution. This will be done piece by piece. In particular, the following functions will be written and used:

* $\texttt{is_integer()}$
* $\texttt{is_positive()}$
* $\texttt{is_nonnegative()}$
* $\texttt{is_positive_integer()}$
* $\texttt{is_nonneg_integer()}$
* $\texttt{is_probability()}$
* $\texttt{bin_factorial()$
* $\texttt{bin_combinations()}$
* $\texttt{bin_probability()}$
* $\texttt{bin_distribution()}$
$\texttt{bin_probability()}$ and $\texttt{bin_distribution()}$ are high level functions that will use the other functions as auxillary functions.

### $\texttt{is_integer()}$
This first function tests whether a given input is an integer or not

```{r}
#' @title Integer test function
#' @description This function tests whether a 
#' given number is an integer value. If the value in
#' question is an integer, then the function will
#' return TRUE. If it is not, FALSE will be returned
#' @param x. numerical value (numeric)
#' @return y. Truth value about x (BOOL)

is_integer <- function(x){
  if(x %% 1 == 0){
    return(TRUE)
  } else{
    return(FALSE)
  }
}
```
Testing the functionality of $\texttt{is_integer()}$
```{r}
# TRUE's
is_integer(-1)
is_integer(0)
is_integer(2L)
is_integer(2)
# FALSE's
is_integer(2.1)
is_integer(pi)
is_integer(0.01)

```

### $\texttt{is_positive()}$
This function tests wether a numeric value is positive or negative. A boolean truth value will be returned

```{r}

#' @title Positivity checker
#' @description Checks whether a numerical value is 
#' greater than zero or less than zero.
#' @param x. numerical value in question (numeric)
#' @param y. Truth value about x (BOOL)

is_positive <- function(x){
  if(x > 0){
    return(TRUE)
  } else{
    return(FALSE)
  }
}
```
Testing the function
```{r}
# TRUE's
is_positive(0.01)
is_positive(2)
# FALSE's
is_positive(-2)
is_positive(0)
```
### $\texttt{is_nonnegative()}$
This function tests whether a number is a non-negative value. It returns a boolean truth value

```{r}
#' @title Inclusive positivity checker
#' @description This function checks whether a 
#' numerical value is non-negative. This differs
#' from the function is_positive in that the value 
#' 0 evaluates to TRUE instead of FALSE
#' @param x. Numerical value in questions (numeric)
#' @param y. Truth value about x (BOOL)

is_nonnegative <- function(x){
  if(x >= 0){
    return(TRUE)
  } else{
    return(FALSE)
  }
}
```
Testing the function
```{r}
# TRUE's
is_nonnegative(0)
is_nonnegative(2)
# FALSE's
is_nonnegative(-0.00001)
is_nonnegative(-2)
```


### $\texttt{is_positive_integer()}$
This function uses $\texttt{is_positive()}$ and $\texttt{is_integer()}$ to check if a numerical value is an integer and if it is non negative. The function returns boolean truth values
```{r}
#' @title Positive integer checker
#' @description Checks whether a given numerical value
#' is both an integer and positive. The function
#' uses the functions is_positve and is_integer in 
#' establishing the truth value of the numerical
#' value in questions
#' @param x. The numerical value in question (numeric)
#' @param y. A truth value of whether x is both an 
#' integer and positive (BOOL)

is_positive_integer <- function(x){
  if(is_integer(x) == TRUE & is_positive(x) == TRUE){
    return(TRUE)
  } else{
    return(FALSE)
  }
}
```
testing the function
```{r}
# TRUE
is_positive_integer(2)
is_positive_integer(2L)
# FALSE
is_positive_integer(0)
is_positive_integer(-2)
```
### $\texttt{is_nonneg_integer()}$
This function uses $\texttt{is_nonnegative()}$and $\texttt{is_integer()}$ to check if a numerical value is non-negative and is an integer. 

```{r}
#'@title Nonnegative integer checker
#'@description Checks whether a numerical value
#'is both nonnegative and an integer. Utilizes the
#'functions is_nonnegative and is_integer
#'@param x. Numerical value in question (numeric)
#'@return y. Truth value of x (BOOL)

is_nonneg_integer <- function(x){
  if(is_nonnegative(x) == TRUE & is_integer(x) == TRUE){
    return(TRUE)
  }
  else{
    return(FALSE)
  }
}
```
testing the function
```{r}
# TRUE's
is_nonneg_integer(0)
is_nonneg_integer(1)
# FALSE
is_nonneg_integer(-1)
is_nonneg_integer(-2.5)
```


### $\texttt{is_probability()}$
Checks wheter a given number falls into the interval [0,1]. The function will return a boolean truth value.
```{r}
#' @title Valid probability value checker
#' @description This subroutine checks whether a 
#' given value is a valid probability value. In 
#' particular, it checks whether a value is inside of
#' the interval [0,1]
#' @param x. Numerical value in question (numeric)
#' @return y. Truth value of x (BOOL)

is_probability <- function(x){
  if(x >= 0 & x <= 1){
    return(TRUE)
  } else{
    return(FALSE)
  }
}
```

testing the function
```{r}
# TRUE's
is_probability(0)
is_probability(0.5)
is_probability(1)
# FALSE's
is_probability(-1)
is_probability(1.0000001)
```

### $\texttt{bin_factorial()}$
Calculates the factorial of a given integer.
```{r}
#' @title Factorial function
#' @description Calculates the factorial of a given integer
#' The factorial of a function is given as n! = n(n-1)(n-2)...1
#' @param x. An integer value (numeric)
#' @return y. The factorial of x (numeric)

bin_factorial <- function(x){
  if(x == 0){
    return(1)
  } else{
  total = 1
  for(i in 1:x){
    total <- total * i
    }
  return(total)
  }
}
```
testing the function
```{r}
# valid
bin_factorial(5)
## [1] 120
bin_factorial(0)
## [1] 1
```

### $\texttt{bin_combinations()}$
Using the $\texttt{factorial()}$ function, we will calculate the number of combinations in which $k$ successes can occur in $n$ trials.

```{r}
#' @title n choose k function
#' @description Calculates the values of k 
#' successes in n trials. This function calls upon
#' the function bin_factorial()
#' @param k. number of successes (numeric)
#' @param n. number of possible trials (numeric)
#' @return y. number of k successes that can occure
#' in n possible trials (numeric)

bin_combinations <- function(n,k){
  return(bin_factorial(n)/(bin_factorial(k) * bin_factorial(n-k)))
}
```
testing the function
```{r}
bin_combinations(n = 5, k = 2)
bin_combinations(10, 3)
bin_combinations(4, 4)
```
### $\texttt{bin_probability()}$
Utilizing the functions $\texttt{is_nonneg_integer()}$, $\texttt{is_probability()}$, and $\texttt{bin_combinations()}$ we will calculate the binomial probability
```{r}

#' @title Binomial Probability
#' @description Calculates the probability of
#' getting k successes in n trials
#' @param n. Number of trials (numeric)
#' @param k. Number of successes (numeric)
#' @param prob. Probability of success (numeric)
#' @return y. Probability of getting the k successes (numeric)

bin_probability <- function(trials, success, prob){
  if(success > trials){
    stop("Invalid number of trials")
  }
  else if(is_nonneg_integer(trials) == FALSE | is_nonneg_integer(success) == FALSE){
    if(is_nonneg_integer(trials) == FALSE){
      stop("Invalid trials value")
    }else{
      stop("Invalid success value")
    }
  }
  else if(is_probability(prob) == FALSE){
        stop("Invalid probability value")
      }
  else{
    return(bin_combinations(trials, success) * (prob^success) * ((1-prob)^(trials-success)))
  }
}

```
testing the function
```{r}
# bad trials
bin_probability(trials = 0, success = 2, prob = 0.5)
# bad success
bin_probability(trials = 5, success = 2.5, prob = 0.5)
# success > trials
bin_probability(trials = 5, success = 6, prob = 0.5)
# bad prob
bin_probability(trials = 5, success = 2, prob = -0.5)
# 55 heads in 100 tosses of a loaded coin with 45% chance of heads
bin_probability(trials = 100, success = 55, prob = 0.45)
```
### $\texttt{bin_distribution()}$
Finally, we will look at a distribution of values. One gives a number of trails, and the probability values of possible successful values will be returned as a dataframe.
```{r}
#'@title Binomial distribution
#'@description creates a distribution, given
#'a number of trials and a probability value,
#'of different success rates
#'@param trials. number of trials (numeric)
#'@param prob. Probability of success for a trial (numeric)
#'@return dataframe of distributions (list)

bin_distribution <- function(trials, prob){
  probability <- rep(0,trials+1)
  success <- seq(0,trials,1)
  for( i in 0:(length(success)-1)){
    probability[i+1] <- bin_probability(trials, i, prob)
  }
  return(data.frame("success" = success, "probability" = probability))
}

# binomial probability distribution
bin_distribution(trials = 5, prob = 0.5)
```
